Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/metalang.tex
checkpoint

In this Section, we introduce the $\metaprot$ language which includes
structured data and function definitions, which are sufficiently
expressive to define composable protocol elements such as garbled
gates. The $\metaprot$ language is a \emph{metalanguage}, in the sense
that it produces $\fedprot$ protocols as a result of computation. That
is, $\metaprot$ is a high-level language that generates low-level
protocol code.

@@

\begin{fpfig}[t]{Type Syntax of $\metaprot$ (T) and selected type judgement rules.}{fig-metaprot-types}
{\small
  $$
  \begin{array}{rcl@{\hspace{2mm}}r}
    \srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst & \gdesc{types}\\ [1mm]
    %&&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
    \viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
    \Gamma &::=& \Gamma; x : \tau \mid \varnothing & \gdesc{type environments}    
  \end{array}
  $$
}

@@

Top-level type judgements are of the form $\Gamma \vdash \codebase, e
: \tau, V$, where all the functions in $\codebase$ are well-typed in
$\Gamma$, the top level view effect $V$ is a set of concrete
$\fedprot$ views which are constructed by a disjoint union of the views
in the effect of $e$- the disjointness requirement guarantees that
views are uniquely defined. Our $\metaprot$ type safety result is
formulated as follows. In addition to safe execution of the
metaprogram, it also guarantees safety of the residual $\fedprot$
program. Of course, it is important to note that ``safety'' here
means the usual type safety property that programs do not get stuck,
it does not imply any security hyperproperties. 
\begin{theorem}[$\metaprot$ Type Safety]
  \label{theorem-metalang-safety}
  Given $\codebase$, $e$, and $\Gamma$ with $\Gamma \vdash \codebase,e : \unity : V$,
  then $\config{\varnothing}{e} \redxs \config{\prog}{\ttt{()}}$ where
  $\prog$ is safe with $\iov(\prog) = S \cup V$ for some $S$.
\end{theorem}

>>>>

In this Section, we introduce the $\metaprot$ language which includes
structured data and function definitions, which are sufficiently
expressive to define composable protocol elements such as garbled
gates. The $\metaprot$ language is a \emph{metalanguage}, in the sense
that it produces $\fedprot$ protocols as a result of computation. That
is, $\metaprot$ is a high-level language that generates low-level
protocol code. We endow $\metaprot$ with a type system that enjoys a
type safety result guaranteeing safety of generated protocols. We
consider library implementations for GMW and YGC circuits as
examples. Additional examples are provided in Appendix
\ref{section-examples-lang}.

@@

\begin{fpfig}[t]{Type syntax of $\metaprot$ (T) and selected type judgement rules (B).}{fig-metaprot-types}
{\small
  $$
  \begin{array}{rcl@{\hspace{2mm}}r}
    \srct &::=& \cidty{e} \mid \stringty{e} \mid \bet{e} \mid  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst & \gdesc{types}\\ [1mm]
    %&&  \{ \flab : \srct;\ldots;\flab : \srct \} \mid \tau * \cdots * \tau \rightarrow \tau,\viewst \\[1mm]
    \viewst  &::=& \view{e}{e};\viewst \mid \varnothing   & \gdesc{view effects}\\[1mm]
    \Gamma &::=& \Gamma; x : \tau \mid \varnothing & \gdesc{type environments}    
  \end{array}
  $$
}

@@

Top-level type judgements are of the form $\Gamma \vdash \codebase, e
: \tau, V$, where all the functions in $\codebase$ are well-typed in
$\Gamma$, the top level view effect $V$ is a set of concrete
$\fedprot$ views which are constructed by a disjoint union of the views
in the effect of $e$- the disjointness requirement guarantees that
views are uniquely defined. Our $\metaprot$ type safety result is
formulated as follows. In addition to safe execution of the
metaprogram, it also guarantees safety of the residual $\fedprot$
program (Definition \ref{definition-safety}).
\begin{theorem}[$\metaprot$ Type Safety]
  \label{theorem-metalang-safety}
  Given $\codebase$, $e$, and $\Gamma$ with $\Gamma \vdash \codebase,e : \unity : V$,
  then $\config{\varnothing}{e} \redxs \config{\prog}{\ttt{()}}$ where
  $\prog$ is safe with $\iov(\prog) = S \cup V$ for some $S$.
\end{theorem}

### 2023-07-11 popl24/minicat.tex
checkpoint

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips (including invocations of the oracle)
respectively. Given a program $\prog$, we write $\iov(\prog)$ to
denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
and $V$ is the set of views in $\prog$, we write $\flips(\prog)$
to denote the set $F$ of flip variables in $\prog$, and we
write $\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set of
variables $X$ and parties $P$, we write $X_P$ to denote the subset of
$X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to denote
the subsets belonging to honest and corrupt parties, respectively.

@@

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$. We define a
program as \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
exists $\store'$ where $\config{\store}{\prog} \redxs
\config{\store'}{\varnothing}$. Unsafe programs are those where views
are used before they're defined, or where expressions mix owners of
variables. As a sanity condition, we will only consider safe 
programs in our presentation of $\fedprot$, though in Section
\ref{section-metalang} we define a type system that guarantees
program safety by construction, as demonstrated in Theorem
\ref{theorem-metalang-safety}.

>>>>

We let $x$ range over variables, and will use $s$, $v$, and $f$ to
range over individual secrets, views, and flips respecitively. We let
$X$ range over sets of variables and $S,V,F$ to range over sets of
secrets, views, and flips (including invocations of the oracle)
respectively. Given a program $\prog$, we write $\iov(\prog)$ to
denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
and $V$ is the set of views in $\prog$, we write $\flips(\prog)$ to
denote the set $F$ of flip variables in $\prog$, and we write
$\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set
of variables $X$ and parties $P$, we write $X_P$ to denote the subset
of $X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to
denote the subsets belonging to honest and corrupt parties,
respectively.

@@

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$. We define a
notion of protocol safety as follows, which is a well-formedness
property.
\begin{definition}[Protocol Safety]
  \label{definition-safety}
A program is \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
exists $\store'$ where $\config{\store}{\prog} \redxs
\config{\store'}{\varnothing}$.
\end{definition}
Observe that unsafe programs are those where views
are used before they're defined, or where expressions mix owners of
variables. As a sanity condition, we will only consider safe 
programs in our presentation of $\fedprot$, though in Section
\ref{section-metalang} we define a type system that guarantees
program safety by construction, as demonstrated in Theorem
\ref{theorem-metalang-safety}.
